Abstraction Refinement for Emptiness Checking of Alternating Data Automata

### 2017-10-20 tacas/draft.tex
ok

We have implemented both Algorithm \ref{alg:predabs} and
\ref{alg:impact} in a prototype tool\footnote{The implementation is
  available at \url{https://github.com/cathiec/AltImpact}} that uses
the Z3 SMT solver\footnote{\url{https://github.com/Z3Prover/z3}} for
the satisfiability queries and interpolant generation, in the theory
of linear integer arithmetic (LIA) combined with booleans. We compared
both algorithms with a previous implementation of a trace inclusion
procedure, called
\textsc{Includer}\footnote{\url{http://www.fit.vutbr.cz/research/groups/verifit/tools/includer/}},
that uses on-the-fly determinisation and lazy predicate abstraction
with interpolant-based refinement \cite{Tacas16} in the LIA theory,
without booleans.

The results of the experiments are given in Table
\ref{tab:experiments}. We applied the tool first to several array
logic entailments, which occur as verification conditions for
imperative programs with arrays \cite{cav09} (array\_shift,
array\_simple, array\_rotation 1+2) available online
\cite{ntslib}. Next, we applied it on proving safety properties of
hardware circuits (hw1+2) \cite{smrcka}. Finally, we considered two
timed communication protocols, consisting of systems that are
asynchronous compositions of timed automata, whom correctness
specifications are given by timed automata monitors: a timed version of
the Alternating Bit Protocol (abp) \cite{abp} and a~controller of a
railroad crossing (train) \cite{henzinger:RealTimeSystems}. All
results were obtained on an Intel(R) Core(TM) i7-4650U CPU @ 1.70GHz
with 8GB of RAM. The automata sizes are in bytes and the execution
times are in seconds.

@@

As in the non-deterministic case \cite{mcmillan06}, \impact~
outperforms lazy predicate abstraction for checking emptiness by at
least one order of magnitude. However, both our implementations are
slower than \textsc{Includer}, on average (except for the abp
example). The reason for this is currently under investigation, one
possible bottleneck being the hardness of the combined (LIA+booleans)
interpolation problems, as opposed to converting the entire formula
into DNF, eliminating the boolean variables and using interpolation in
the pure LIA theory.

>>>>

We have implemented both Algorithm \ref{alg:predabs} and
\ref{alg:impact} in a prototype tool\footnote{The implementation is
  available at \url{https://github.com/cathiec/JAltImpact}} that uses
the MathSAT5 SMT solver\footnote{\url{http://mathsat.fbk.eu/}} via the
Java SMT
interface\footnote{\url{https://github.com/sosy-lab/java-smt}} for the
satisfiability queries and interpolant generation, in the theory of
linear integer arithmetic (LIA) combined with booleans. We compared
both algorithms with a previous implementation of a trace inclusion
procedure, called
\textsc{Includer}\footnote{\url{http://www.fit.vutbr.cz/research/groups/verifit/tools/includer/}},
that uses on-the-fly determinisation and lazy predicate abstraction
with interpolant-based refinement \cite{Tacas16} in the LIA theory,
without booleans.

The results of the experiments are given in Table
\ref{tab:experiments}. We applied the tool first to several array
logic entailments, which occur as verification conditions for
imperative programs with arrays \cite{cav09} (array\_shift,
array\_simple, array\_rotation 1+2) available online
\cite{ntslib}. Next, we applied it on proving safety properties of
hardware circuits (hw1+2) \cite{smrcka}. Finally, we considered two
timed communication protocols, consisting of systems that are
asynchronous compositions of timed automata, whom correctness
specifications are given by timed automata monitors: a timed version
of the Alternating Bit Protocol (abp) \cite{abp} and a~controller of a
railroad crossing (train) \cite{henzinger:RealTimeSystems}. All
results were obtained on x86\_64 Linux Ubuntu virtual machine with 8GB
of RAM. The automata sizes are in bytes and the execution times are in
seconds.

@@

As in the non-deterministic case \cite{mcmillan06}, \impact~
outperforms lazy predicate abstraction for checking emptiness by at
least one order of magnitude. However, both our implementations are
slower than \textsc{Includer}, on average. The reason for this is
currently under investigation, one possible bottleneck being the
hardness of the combined (LIA+booleans) interpolation problems, as
opposed to converting the entire formula into DNF, eliminating the
boolean variables and using interpolation in the pure LIA theory.
